% Do not forget to include Introduction
%---------------------------------------------------------------
%\chapter{Introduction}
% uncomment the following line to create an unnumbered chapter
\chapter*{Introduction}\addcontentsline{toc}{chapter}{Introduction}\markboth{Introduction}{Introduction}
%---------------------------------------------------------------
\setcounter{page}{1}

% The following environment can be used as a mini-introduction for a chapter. Use that any way it pleases you (or comment it out). It can contain, for instance, a summary of the chapter. Or, there can be a quotation.
%\begin{chapterabstract}
%\end{chapterabstract}

Modern high-performance network devices are usually proprietary systems that combine custom hardware, specialized operating systems, and tightly coupled software. 
While these solutions offer high throughput and re\-li\-a\-bi\-li\-ty, they are typically expensive, inflexible, and slower to evolve due to their closed design and development model.
Vector Packet Processing (VPP) is a high-performance network stack that operates at layers 2 to 4 of the ISO/OSI model. 
It was originally developed by Cisco Systems, Inc. (which is a world leader in networking) and open-sourced in 2016 under the Fast Data Project (FD.io), that is part of the Linux Foundation.
VPP brings the ability to perform efficient, high-speed packet processing on common off-the-shelf (COTS) hardware, across a wide range of platforms and operating systems.
Its open and flexible architecture opens the door to a new class of network applications that can be deployed and scaled more easily than traditional hardware appliances. 
In this way, VPP could represent a shift in the traditionally conservative networking world, echoing the ‘‘Mainframe to PC’’ revolution, 
where general-purpose systems replaced proprietary platforms, enabling broader innovation and accessibility.

Since VPP was open-sourced only recently, it has not yet been widely adopted by the market, and there are only a limited number of academic studies on the subject. As a result, this area remains underexplored. 
This thesis aims to contribute to this field by evaluating VPP's\footnote{The abbreviation VPP is also commonly used in academic literature to refer to a Virtual Power Plant.} performance, 
with a focus on comparing throughput, latency, and energy efficiency across different VPP configurations and a reference Linux-based router under various traffic conditions.
The goal is to understand the theoretical advantages of VPP, design and implement a suitable measurement infrastructure, define representative traffic scenarios, 
and ultimately evaluate whether VPP provides superior performance. The aim is also to characterize the specific conditions and configurations under which it outperforms the Linux-based reference system.
The results may support a recommendation for its deployment by small-to-mid-sized ISPs currently using Linux-based routers or planning to upgrade legacy systems.
Moreover, in light of the increasing focus on energy efficiency in IT infrastructure, the results may also inform ongoing discussions about sustainable network architecture.

With the development of AI and the growing demand for high-resolution streaming services, it is highly likely that the demand for internet bandwidth will continue to rise. 
This will result in an increased need for network equipment capable of processing larger volumes of data more efficiently. 
Therefore, it is crucial to explore technologies like VPP that are capable to handle this growing demand and to explore their energy efficiency.

This thesis is divided into two parts: Theoretical and Practical. 
The Theoretical part presents the traditional approach to networking and packet processing, as well as an overview of how VPP is designed and the principles on which it operates. 
Additionally, it introduces the testing scenarios that were used. 
The Practical part describes the testing infrastructure, presents the results of various measurements, and provides an analysis of the findings.

%---------------------------------------------------------------
\chapter{Theoretical part}
%---------------------------------------------------------------

%---------------------------------------------------------------
\section{Vector Packet Processing (VPP) and Its Operating Principles}
%---------------------------------------------------------------
\input{text/chapters/principles.tex}

%---------------------------------------------------------------
\section{Implementation of Vector Packet Processing}
%---------------------------------------------------------------
\input{text/chapters/implementation}

%---------------------------------------------------------------
\section{Utilization of Vector Packet Processing}
%---------------------------------------------------------------
\input{text/chapters/usage.tex}

%---------------------------------------------------------------
\section{Survey of Traffic Generation Tools}
%---------------------------------------------------------------
\input{text/chapters/pktgen.tex}

%---------------------------------------------------------------
\chapter{Pratical part}
%---------------------------------------------------------------

%---------------------------------------------------------------
\section{Building Infrastructure for Measurement}
%---------------------------------------------------------------
\input{text/chapters/infrastructure.tex}

%---------------------------------------------------------------
\section{Metodology}
%---------------------------------------------------------------
\input{text/chapters/metodology.tex}

%---------------------------------------------------------------
\section{Test Scenarios and Results}
%---------------------------------------------------------------
\input{text/chapters/measurement.tex}
\input{text/chapters/measurement2.tex}
\input{text/chapters/measurement3.tex}

%---------------------------------------------------------------

%---------------------------------------------------------------
\section{Presentation and Analysis of Results -> do závěru?}
\input{text/chapters/analysis.tex}
%---------------------------------------------------------------


\chapter{Conclusion and Future work}
The main objective of this thesis was to analyze and compare the performance, latency, and energy efficiency of different VPP configurations with a traditional Linux-based router. 
The work aimed to explore the theoretical advantages of VPP, design and implement a custom measurement infrastructure, define relevant traffic scenarios, 
and determine the conditions under which VPP outperforms the Linux network stack.

The theoretical part of the thesis provided a detailed explanation of VPP's architecture and operating principles. 
It also included a comparative analysis of packet processing between VPP and the traditional Linux network stack, highlighting the key differences in design, performance potential, and resource utilization.

As part of the practical work, a custom measurement infrastructure was designed and implemented to enable reliable and repeatable testing. 
Using this setup, a total of 4800 measurements, reflecting realistic usage conditions on standard COTS hardware, 
were conducted across 60 different test scenarios using four configurations -- three VPP variants and one Linux-based reference system.
This resulted in 240 distinct test cases, providing a robust dataset for evaluating throughput, latency, and energy efficiency under a variety of traffic conditions.

The results show that VPP consistently outperformed the Linux-based reference system in terms of throughput and latency -- with the exception of low-traffic forwarding scenarios with higher frame sizes, 
where the Linux stack achieved slightly better results. 
However, these differences were relatively minor. 
In terms of energy efficiency, the Linux-based reference system was generally more efficient than all VPP configurations, with the exception of VPP-1, which outperformed Linux in all cases.
VPP also demonstrated more stable performance.
Given VPP’s overall higher performance ceiling, it is reasonable to expect that it would handle sudden traffic spikes more effectively and with greater stability compared to the Linux-based solution. 
The results also clearly indicate that VPP would be more resilient under DoS or DDoS-style attack conditions.


Several directions for future work remain open. For small-to-mid-sized ISPs, it would be useful to compare VPP not only with the Linux network stack, but 
also with commercial or legacy routers commonly used in practice. 
It would also be valuable to evaluate the impact of advanced features such as Quality of Service (QoS), firewalling, or traffic shaping. 
Finally, comparing VPP with a custom-tuned Linux kernel optimized for networking could provide deeper insight into the limits of traditional software-based routing.
